---
title: "Are environmental and geographic effective surrogates for genetic variation in conservation planning?"
author: |
  | Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Cynthia Riginos$^2$, Hugh P. Possingham$^1$, Richard A. Fuller$^1$
  | $^1$School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
  | $^2$School of Geography, Planning and Environmental Management, The University of Queensland, Brisbane, QLD, Australia
  | Correspondance should be addressed to jeffrey.hanson@uqconnect.edu.au
date: "`r format(Sys.time(), '%d %B %Y')`"
tags: ["conservation", "biodiversity", "AFLP", "reserve-selection"]
abstract: "Insert abstract here."
output:
  rmarkdown::pdf_document:
    toc: true
    toc_depth: 5
    keep_tex: yes
    fig_caption: yes
    includes:
      in_header: preamble-latex.tex
fontsize: 11pt
documentclass: article
bibliography: Endnote_lib.bib
csl: reference-style.csl
vignette: >
  %\VignetteIndexEntry{rapr}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, iclude=FALSE}
# set global cache
knitr::opts_chunk$set(cache=FALSE)

# set seed for reproducibility
set.seed(500)
```
 
## Introduction


## Methods


### Study area
To address the aims of this study, we obtained species distribution and genomic (AFLP) from [@r451]. This dataset was chosen because it provides genomic data for a multitude of species at a high spatial resolution (approx. $20$km$^{2} \times 22$km$^{2}$).


```{r}
## compile spatial data
# load grid cells
grid.DF <- fread(
	'inst/extdata/Data Meirmans et al IntrabioDiv/ReadMe.txt',
	data.table=FALSE,
	skip='cell\tLong\tLat'
) %>% rename(
		grid.longitude=Long,
		grid.latitude=Lat
)

# load in distribution data
spp.loc.files <- dir(
	'inst/extdata/Data Meirmans et al IntrabioDiv',
	'^.*locations\\.txt$',
	full.names=TRUE)

spp.loc.DF <- ldply(
	spp.loc.files,
	.fun=function(x) {
		mutate(
			fread(x, data.table=FALSE),
			species=gsub('_', '', gsub('_locations.txt', '', basename(x), fixed=TRUE), fixed=TRUE),
			presence='present'
		) %>% rename(
			cell=population,
			sample.longitude=longitude,
			sample.latitude=latitude
		)
	}
)

# make presence-absence data
spp.loc.DF <- ddply(spp.loc.DF, 'species', .fun=function(x) {
	# merge presences with grid data
	x %>% full_join(
		grid.DF,
		by='cell',
		all=TRUE
	) %>% mutate(
		species=replace(species, is.na(species),
			first(na.omit(species))),
		presence=replace(presence, is.na(presence),
			'absent')
	) %>% arrange(
		cell
	)
})

```

```{r}
## plot map of species distributions'
# download basemap
data(countriesHigh)
countries.PL <- countriesHigh[
	countriesHigh$ADMIN %in% c(
		'Italy', 'Switzerland', 'France', 'Austria',
		'Germany', 'Slovenia', 'Croatia', 'Hungary',
		'Monaco', 'Germany'
	)
,]

# plot species data
ggplot() +
	geom_polygon(data=countries.PL, aes(x=long, y=lat, group=group)) +
	geom_point(
		data=spp.loc.DF, 
		aes(x=grid.longitude, y=grid.latitude, color=presence),
		size=2.5
	) +
	scale_color_manual(
		name='presence',
		values=c('present'='red', 'absent'='transparent')
	) +
	theme_classic() +
	theme(axis.ticks=element_blank(), axis.text=element_blank()) +
	coord_cartesian(
		xlim=buffered.range(spp.loc.DF$grid.longitude, 0.05),
		ylim=buffered.range(spp.loc.DF$grid.latitude, 0.05)
	) +
	xlab('') +
	ylab('') +
	facet_wrap(~ species, ncol=5)

```

```{r}
# calculate species richness for each planning unit
rich.loc.DF <- spp.loc.DF %>% filter (
	presence=='present'
) %>% group_by(
	cell
) %>% summarise(
	grid.longitude=first(grid.longitude),
	grid.latitude=first(grid.latitude),
	richness=n_distinct(species)
)

# plot map of species richness
ggplot() +
	geom_polygon(data=countries.PL, aes(x=long, y=lat, group=group), color='gray20') +
	geom_point(
		data=rich.loc.DF, 
		aes(x=grid.longitude, y=grid.latitude, color=richness),
		size=4
	) +
	theme_classic() +
	theme(axis.ticks=element_blank(), axis.text=element_blank()) +
	coord_cartesian(
		xlim=buffered.range(spp.loc.DF$grid.longitude, 0.05),
		ylim=buffered.range(spp.loc.DF$grid.latitude, 0.05)
	) +
	xlab('') +
	ylab('')

```


### Genomic data


### Surrogate data

```{r}
# compile spatial data

```

```{r}
# plot map of averages per quadrat

```

### Prioritisations



## Results

## Discussion

##

## Acknowledgements
JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF has an Australian Research Council Future Fellowship. This work was supported by the Centre of Excellence for Environmental Decisions (CEED) and the Landscape Ecology and Conservation Group (LEC) at The University of Queensland.

## References
<span style="color:white">blank</span>
